Online Sequential Monte Carlo smoother for partially observed diffusion processes

### 2017-08-25 TexSubmission/ParisEM.tex
parenthesis

\begin{figure}[p]
\centering
\includegraphics[scale=0.5]{res_mult_SINE_cv}
\caption{{\em SINE model - variance}. Distribution of the empirical absolute coefficient of variation.}
\label{fig:mult:SINE:cv}
\end{figure}
\paragraph{Generalized EM procedure}
The performance of our algorithm is also assessed in the case where $\theta$ and the variance $\sigma^2_\text{obs}$ are unknown and estimated using a generalized EM algorithm. 
The study is done using a data set with $n = 200$ observations simulated with $\mu=0$ an d $ \sigma^2_\text{obs} = 1$. The Grand PaRIS algorithm is used to perform the E step, with the same settings as before for $N$, $\tilde{N}$ and $M$.
As their is no closed form solution to compute the M step of the EM algorithm and propose new parameter estimates, we use a Generalized EM procedure: given the current estimation $\theta^{(k)} := (\mu^{(k)}, \sigma_{\text{obs}}^{(k)}$, the function $\mathcal{Q}(\cdot, \theta^{(k)})$ is approximated for $50$ new candidates $\theta_1, \dots, \theta_{50} $ chosen by the user. The new estimate is set as
\[
\theta^{(k + 1)} = \text{argmax}_i\mathcal{Q}(\theta_i, \theta^{(k)})\eqsp .
\]
This procedure has the nice property of using the same particle filter and the same retrospective sampling of Lemma \ref{lem:AR:unbiased} for all candidates, avoiding to repeat this time consuming procedure. 
The number of candidates and the way to choose them is problem dependent and then left to the user. 
In our case, we sampled candidates using Gaussian distributions around the currend estimate $\theta^{(k)}$, decreasing the variance when $k$ increases.
Figures~\ref{fig:SINE:theta} and~\ref{fig:SINE:sigma} illustrate the performance of the estimation for 12 different initializations of $\mu$ (resp. $\sigma_\text{obs}$) uniformly chosen in  $(-\pi,\pi)$ (resp. in $(0,6)$), showing a convergence after only a few iterations of the EM procedure.

>>>>

\begin{figure}[p]
\centering
\includegraphics[scale=0.5]{res_mult_SINE_cv}
\caption{{\em SINE model - variance}. Distribution of the empirical absolute coefficient of variation.}
\label{fig:mult:SINE:cv}
\end{figure}
\paragraph{Generalized EM procedure}
The performance of our algorithm is also assessed in the case where $\theta$ and the variance $\sigma^2_\text{obs}$ are unknown and estimated using a generalized EM algorithm. 
The study is done using a data set with $n = 200$ observations simulated with $\mu=0$ an d $ \sigma^2_\text{obs} = 1$. The Grand PaRIS algorithm is used to perform the E step, with the same settings as before for $N$, $\tilde{N}$ and $M$.
As their is no closed form solution to compute the M step of the EM algorithm and propose new parameter estimates, we use a Generalized EM procedure: given the current estimation $\theta^{(k)} := \left(\mu^{(k)}, \sigma_{\text{obs}}^{(k)} \right)$, the function $\mathcal{Q}(\cdot, \theta^{(k)})$ is approximated for $50$ new candidates $\theta_1, \dots, \theta_{50} $ chosen by the user. The new estimate is set as
\[
\theta^{(k + 1)} = \text{argmax}_i\mathcal{Q}(\theta_i, \theta^{(k)})\eqsp .
\]
This procedure has the nice property of using the same particle filter and the same retrospective sampling of Lemma \ref{lem:AR:unbiased} for all candidates, avoiding to repeat this time consuming procedure. 
The number of candidates and the way to choose them is problem dependent and then left to the user. 
In our case, we sampled candidates using Gaussian distributions around the currend estimate $\theta^{(k)}$, decreasing the variance when $k$ increases.
Figures~\ref{fig:SINE:theta} and~\ref{fig:SINE:sigma} illustrate the performance of the estimation for 12 different initializations of $\mu$ (resp. $\sigma_\text{obs}$) uniformly chosen in  $(-\pi,\pi)$ (resp. in $(0,6)$), showing a convergence after only a few iterations of the EM procedure.
